{\fontfamily{ptm}\selectfont Reading Notes}

### 2025-05-29 Notebook LaTEX/Formalized Notes/Notes 2025 03/main.tex
>w<, changes

\chapter{Advanced Enhancement Techniques in Accelerated Proximal Gradient}
    We review advanced enhancement techniques in Accelerated Proximal Gradient method. 
    The review will be based on several papers. 
    \par
    There are several notable enhancements of the FISTA for function that are not strongly convex. 
    Monotone variants of FISTA proposed by Beck \cite{beck_fast_2009-1} and Nesterov \cite[2.2.32]{nesterov_lectures_2018} imposes monotonicity in function value at the iterates.  
    Backtracking strategies from Chambolle \cite{calatroni_backtracking_2019} shows that the underestimating Lipschitz constant using a backtracking technique to choose a next iterate improves the average runtime of the algorithm in practice. 
    They showed that the convergence rate is bounded by the estimates of the Lipschitz constant. 
    Restart is a technique discussed in Necoara et al. \cite{necoara_linear_2019}.
    They showed that there exists an optimal restarting interval to achieve fast line convergence rate for all functions with quadratic growth condition. 
    The method of restarting was improved to be parameter free while still retaining a fast linear convergence rate, see Alamo et al. \cite{alamo_restart_2019} and Aujol et al. \cite{aujol_parameter-free_2024}. 
    \par
    In this chapter, we will go through the details of these enhancements of FISTA and discuss why they are important in theories, and in practice.

@@

                If a function is $L$ smooth, it also be $L$ weakly convex. 

@@

                And descent is possible when $\beta \ge (L + q_g)/2$ and it yields the descent lemma: 
                    0 &\le F(x) - F(\bar x) - 1/\beta \Vert \mathcal G_{1/\beta}(x)\Vert^2. 
                The maximum amount of descent is achieved when $\beta = L + q_g$. 

@@

            % Suppose that the sequence $(\alpha_k)_{k \ge 0}$ has $\alpha_0 \in (0, 1]$ and for all $k \ge 1$, it satisfies $\alpha_{k - 1}^2\rho_{k - 1}(1 - \alpha_k) = \alpha_k^2$. 
            % Suppose that $(\beta_k)_{k \ge 0}$ is a sequence such that $\beta_0 = 1$ and for all $k \ge 1$ it's defined as: 

@@

                \item[(b)] Because $\rho_{k}L_{k + 1}L_k = 1$ for all $k \ge 0$, the definition of $(\beta_k)_{k\ge 0}$ simplifies and $\beta_k = (\alpha_k^2/\alpha_0^2)(L_k/L_0)$. As a consequence it also gives for all $k \ge 1$ that: 

@@

        The following assumption will characterize the full scope of discussion in nonconvexity objective function. 

@@

            \STATE{$n_1:=p_0. $}
            \STATE{$M := M - n_1$}

@@

        \begin{observation}

@@

            For all $J -1 \ge j \ge 1 $ it has $p_{j-1}\le p_{j}$ so $(p_j)_{j \ge 1}^{J}$ is monotone exluding the last iterates $p_J$. 

@@

        When $j = J$ it's not necessarily true that $p_J \ge n_J$ because on the last iteration, it can exit the forloop through conditions $G_k^+ \le \epsilon$ in line 14 of Algorithm \ref{alg:gmapg}, but $n_J \ge p_{J - 1} \ge n_{J - 1}$ still. 

@@

        So $x_{p_j}$ is assigned to assigned to $z_{j + 1}$ by the outer loop and it has $x_{p_j} = x_{p_j}^{(j)} = z_{j + 1}$. 

@@

            Suppose that $F = f + g$ satisfies Assumption \ref{ass:standard-fista} and Assumption \ref{ass:q-growth-ch2}, denote $F^+ := \inf_xF(x)$ and, $X^+$ as the set of minimizers. 

@@

        \begin{lemma}[bounds on outer iteration counts]

@@

            for $j \le t \le j + T_\epsilon - 1$, and denote $s = j + T_\epsilon -1$ for better notations, giving us $j \le t \le s$. 

@@

        \end{proof}
        \begin{theorem}[convergence rates of restarted GMAPG]

>>>>

\chapter{Advanced Enhancement Techniques in Accelerated Proximal Gradient}
    We review advanced enhancement techniques in Accelerated Proximal Gradient method. 
    The review will be based on several papers. 
    \par
    There are several notable enhancements of the FISTA for function that are not strongly convex. 
    Monotone variants of FISTA proposed by Beck \cite{beck_fast_2009-1} and Nesterov \cite[2.2.32]{nesterov_lectures_2018} imposes monotonicity in function value at the iterates.  
    Backtracking strategies from Chambolle \cite{calatroni_backtracking_2019} shows that the underestimating Lipschitz constant using a backtracking technique to choose a next iterate improves the average runtime of the algorithm in practice. 
    They showed that the convergence rate is bounded by the estimates of the Lipschitz constant. 
    Restart is a technique discussed in Necoara et al. \cite{necoara_linear_2019}.
    They showed that there exists an optimal restarting interval to achieve fast linear convergence rate for all functions with quadratic growth condition. 
    The method of restarting was improved to be parameter free while still retaining a fast linear convergence rate, see Alamo et al. \cite{alamo_restart_2019} and Aujol et al. \cite{aujol_parameter-free_2024}. 
    \par
    In this chapter, we will go through the details of these enhancements of FISTA and discuss why they are important in theories, and in practice.

@@

                If a function is $L$ smooth, it's $L$ weakly convex also. 

@@

                And descent is possible when $\beta \ge (L + q_g)/2$ and, it yields the descent lemma: 
                    F(\bar x) - F(x) &\le - 1/\beta \Vert \mathcal G_{1/\beta}(x)\Vert^2. 

@@

@@

                \item[(b)] Because $\rho_{k}L_{k + 1}L_k^{-1} = 1$ for all $k \ge 0$, the definition of $(\beta_k)_{k\ge 0}$ simplifies and $\beta_k = (\alpha_k^2/\alpha_0^2)(L_k/L_0)$. As a consequence it also gives for all $k \ge 1$ that: 

@@

        The following assumption will characterize the full scope of discussion in nonconvex objective function. 

@@

            \STATE{$n_1:=p_0$.}
            \STATE{$M := M - n_1$.}

@@

        \begin{observation}\label{obs:rgmapg}

@@

            For all $J -1 \ge j \ge 1 $ it has $p_{j-1}\le p_{j}$ so $(p_j)_{j \ge 1}^{J}$ is monotone excluding the last iterates $p_J$. 

@@

        When $j = J$ it's not necessarily true that $p_J \ge n_J$ because on the last iteration, it can exit the for loop through conditions $G_k^+ \le \epsilon$ in line 14 of Algorithm \ref{alg:gmapg}, but $n_J \ge p_{J - 1} \ge n_{J - 1}$ still. 

@@

        So $x_{p_j}$ is assigned to $z_{j + 1}$ by the outer loop and it has $x_{p_j} = x_{p_j}^{(j)} = z_{j + 1}$. 

@@

            Suppose that $F = f + g$ satisfies Assumption \ref{ass:q-growth-ch2}, denote $F^+ := \inf_xF(x)$ and, $X^+$ as the set of minimizers. 

@@

        \begin{lemma}[bounds on outer iteration counts]\label{lemma:rgmapg-outer-itr-bnd}
            Let $F = f + g$ satisfies Assumption \ref{ass:standard-fista}. 
            Suppose we apply Algorithm \ref{alg:linear-rgmapg} on $F = f + g$. 
            Define $T_\epsilon = \lceil\ln(2\epsilon^{-2}(F(z_0) - F^+))\rceil$
            Assume that after iteration $j$, no doubling occurred in the if statement, i.e: $n_{t + 1} = p_t$ for $t \ge j$, then it must terminate before, or at iteration $j + T_\epsilon$. 

@@

            for $j \le t \le j + T_\epsilon - 1$, and denote $s = j + T_\epsilon -1$ for better notations, giving us for all $j \le t \le s$, $n_{t + 1} = p_t$. 
            Consider the start of the $s$ th iteration of the outer loop, denote $L_0^{(s)}, G_0^{(s)}$ as the $L_0$ in line 4 in Algorithm \ref{alg:gmapg}, it would give the following inequalities
                \frac{1}{2}\left(
                    G_0^{(s)}
                \right)^2
                &\underset{\text{(a)}}{=} 
                \frac{1}{2}\left\Vert
                    \sqrt{L_0^{(s)}}\left(z_s - T_{1/L_0^{(s)}}(z_s)\right)
                &\underset{\text{(b)}}{=} \frac{1}{2L_0^{(s)}}\left\Vert
                        \mathcal G_{1/L_0^{(s)}}(z_s)
                &\underset{\text{(c)}}{\le} F(z_s) - F\left(x_0^{(s)}\right) 
                &\underset{\text{(d)}}{\le} 
                F(z_s) - F\left(x_{p_s}^{(s)}\right)
                &= F(z_s) - F(z_{s + 1})
                &\underset{\text{(e)}}{\le} \exp(-T_\epsilon)\left(F(z_{s - T_\epsilon}) - F(z_{s - T_\epsilon + 1})\right)
                &= \exp(-T_\epsilon)(F(z_{j - 1}) - F(z_j)) 
                &\underset{\text{(f)}}{\le} \exp(-T_\epsilon)(F(z_{0}) - F(z_j))
                &\underset{\text{(g)}}{\le} \left(
                    \frac{2(F(z_0) - F^+)}{\epsilon^2}
                \right)^{-1}(F(z_0) - F^+) 
                &= \epsilon^2/2.
                \item [(a)] At the $s$ iteration of the loop in Algorithm \ref{alg:linear-rgmapg}, $z_s$ is passed into Algorithm \ref{alg:gmapg} with $x_{-1} = z_s$. Therefore, at line 3 in Algorithm \ref{alg:gmapg} it calls Algorithm \ref{alg:armijo-ls} with $x_{-1} = z_s, \alpha_0 = 1, L^+$ which means $y^+ = x_{-1} = z_s$ at line 5 and $x^+ = T_{1/L^+}(z_s)$. Coming back to line 3 in Algorithm \ref{alg:gmapg}, it assigns $y_0 = y^+ = z_s, x_0 = x^+ = T_{1/L^+}(z_s)$ and $L_0 = L^+$. Assuming the line search went successful, it will have $D_f(z_s, x_0) \le L^{(s)}_0/2 \Vert z_s - x_0\Vert^2$. 
                \item [(b)] We used definition of gradient mapping in Definition \ref{def:gm-for-ch2}. 
                \item [(c)] By the assumption that the line search in Algorithm \ref{alg:armijo-ls} is successful at $z_s$ back in item (a), here we can use \eqref{eqn:emp:result-item-2} in Theorem \ref{thm:gmapg-generic-gm-cnvg}. 
                \item [(d)] GMAPG is monotone in function value for the use of $\mathbf M$ that is either Algorithm \ref{alg:beck-mono} or \ref{alg:nes-mono}, so it has $F\left(x_{p_s}^{(s)}\right) \le F\left(x_0^{(s)}\right)$. 
                \item [(e)] Here we used the assumption that no doubling occurs so $n_{t + 1} = p_t$ for all $j \le t \le s$ meaning that line 13 in Algorithm \ref{alg:linear-rgmapg} has $F(z_t) - F(z_{t + 1}) \ge e^{-1}(F(z_{t - 1}) - F(z_t))$. For that, we recursively unrolled it for $T_\epsilon$ many iterations starting with $t = s = j + T_\epsilon - 1$ ending with $t = j$. 
                \item [(f)] We used the monotone property of subroutine $\mathbf M$ in GMAPG again so $F(z_0) \le F(z_{j - 1})$. 
                \item [(g)] We substituted $T_\epsilon = \lceil\ln(2\epsilon^{-2}(F(z_0) - F^+))\rceil$ and, removing $\lceil\cdot\rceil$ to make for the $\le$ inequality. 
            Therefore, it has $G_0^{(s)} \le \epsilon$, hence it must have terminated at, or before iteration $s$. 

@@

            Firstly, Algorithm \ref{alg:linear-rgmapg} must terminate within finite many iterations under Assumption \ref{ass:q-growth-ch2} for all total budget $M \in \N$. 
            This is because Observation \ref{obs:rgmapg} shows that $n_j$ is a non-decreasing sequence, Lemma \ref{lemma:rgmapg-inner-bnds} shows that $n_j \le \bar p$ under this quadratic growth assumption, therefore $n_j$ must converge which implies that doubling of $n_{j + 1} = 2p_{j}$ must stop. 
            Finally, since the doubling must stop, Lemma \ref{lemma:rgmapg-outer-itr-bnd} applies hence, it terminates at most $j + T_\epsilon$ iteration. 
            Using it let's assume that it executed for $j = 1, 2, \cdots, J$, and it can be represented as $J = m + nT_\epsilon$ with $0 \le m < T_\epsilon$. 
            The following intermediate results are important to the proof. 
                \item [(a)] 
        \begin{theorem}[convergence rates of restarted GMAPG]
